---
title: "Replacing String and NA"
author: "JS"
date: "10/12/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Replacing stirng in r

Using library tidyverse and stringr
Import excel file from 
library(readxl)
CSE_Analysis_Full_Data_2020_Cohort_missing_hrs_ap_10_9 <- read_excel("C:/Users/jsparks3/Downloads/CSE Analysis Full Data 2020 Cohort missing hrs_ap 10-9.xlsx")
myup2020 = CSE_Analysis_Full_Data_2020_Cohort_missing_hrs_ap_10_9 

This data included 21 NAs
```{r myup2020}
 library(readxl)
CSE_Analysis_Full_Data_2020_Cohort_missing_hrs_ap_10_9 <- read_excel("CSE Analysis Full Data 2020 Cohort missing hrs_ap 10-9.xlsx",na="")
myup2020 <- CSE_Analysis_Full_Data_2020_Cohort_missing_hrs_ap_10_9
library(Hmisc)
Hmisc::describe(myup2020[, 9:20])
```
## Histogram for GPA high school
```{r}
myup2020$GPA_HIGHSCHOOL <- as.numeric(myup2020$GPA_HIGHSCHOOL)
hist(myup2020$GPA_HIGHSCHOOL)
```

## Including GPA High School Plots

You can also embed plots, for example:

```{r myup2020 plot for High school GPA, echo=FALSE}
library(ggplot2)
library(dplyr)
meanHSGPA1 <- myup2020 %>% 
  group_by(ENTRY_COLLEGE) %>% 
  summarise(meanHSGPA=mean(GPA_HIGHSCHOOL, na.rm = TRUE),CFTIC=n(), .groups="drop")
meanHSGPA1
Countabove4.00 <- myup2020 %>% filter(GPA_HIGHSCHOOL >= 4.00) %>% 
  group_by(ENTRY_COLLEGE) %>% 
  summarise(meanHS =n() );Countabove4.00
ggplot(meanHSGPA1, aes( x=ENTRY_COLLEGE, y=meanHSGPA, col=ENTRY_COLLEGE))+geom_point(size=5)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
##Replace Gender F and M for Female and Male with two differenct ways
### Replace NA for zero values for numerical columns
```{r}
library(stringr)
library(dplyr)
unique(myup2020$GENDER)
myup2020$GENDER <- str_replace_all( myup2020$GENDER,"Female","F")
myup2020$GENDER <- str_replace_all( myup2020$GENDER,"Male","M")
unique(myup2020$GENDER)
myup2020$GENDER <- str_replace_all(myup2020$GENDER, c("F"="Female", "M"="Male"))
unique(myup2020$GENDER)
myup2020$GENDER <- str_replace_na(myup2020$GENDER, "NA")
unique(myup2020$GENDER)
myup2020$GENDER <- ifelse(myup2020$GENDER == "Female", "F", ifelse(myup2020$GENDER=="Male","M", NA))
unique(myup2020$GENDER)
unique(myup2020$GPA_HIGHSCHOOL)
sum(is.na(myup2020$GPA_HIGHSCHOOL))
myup2020$GPA_HIGHSCHOOL[is.na(myup2020$GPA_HIGHSCHOOL)] <- 0
sum(is.na(myup2020$GPA_HIGHSCHOOL))

myup2020NAomit <- myup2020 %>% 
  select(ENTRY_COLLEGE,ENTRY_PROGRAM,GENDER,COUNTY,COUNTY_GROUP,GPA_HIGHSCHOOL,APPLICANT_TIER,HIGH_SCHOOL_NAME,
        ACT_PROPORTION,SAT_PROPORTION)
myup2020NAomit$GPA_HIGHSCHOOL <- na_if(myup2020NAomit$GPA_HIGHSCHOOL,0)
myup2020NAomit$GPA_HIGHSCHOOL <- as.numeric(myup2020NAomit$GPA_HIGHSCHOOL)

unique(myup2020NAomit$APPLICANT_TIER)

sapply(myup2020NAomit, function(x){sum(is.na(x))*100/length(x)})

```
## Na Imputation Methods (Used as regressions and range)
Red plot shows the proportion of NA and yellow shows that your imputation still have some NA and this means
the imputations using variables could not be applied to thoses observations
```{r}
library(mice)
library(VIM)
myup2020NAomit2 <- myup2020NAomit %>% 
  select(GPA_HIGHSCHOOL,APPLICANT_TIER,ACT_PROPORTION) ;glimpse(myup2020NAomit2)
myup2020NAomit2$GPA_HIGHSCHOOL <- as.numeric(gsub("\\.","",myup2020NAomit2$GPA_HIGHSCHOOL))
myup2020NAomit2$APPLICANT_TIER <- as.numeric(gsub("\\.","",myup2020NAomit2$APPLICANT_TIER))
myup2020NAomit2$ACT_PROPORTION <- as.numeric(gsub("\\.","",myup2020NAomit2$ACT_PROPORTION))


md.pattern(myup2020NAomit2)
md.pairs(myup2020NAomit2)
marginplot(myup2020NAomit2[, c("APPLICANT_TIER","GPA_HIGHSCHOOL")])
myup2020NAomit2$APPLICANT_TIER <- log(myup2020NAomit2$APPLICANT_TIER)
myup2020NAomit2$GPA_HIGHSCHOOL <- as.numeric(myup2020NAomit2$GPA_HIGHSCHOOL)
myup2020NAomit2$GPA_HIGHSCHOOL <- log(myup2020NAomit2$GPA_HIGHSCHOOL)
aggr(myup2020NAomit2)
imp_regression <- regressionImp(GPA_HIGHSCHOOL ~ ACT_PROPORTION+APPLICANT_TIER, myup2020NAomit2)
imp_ranger <- rangerImpute(GPA_HIGHSCHOOL~ACT_PROPORTION+APPLICANT_TIER, myup2020NAomit2)
aggr(imp_regression, delimiter = "_imp")
aggr(imp_ranger, delimiter = "_imp")
imp_regression[ , c("GPA_HIGHSCHOOL", "APPLICANT_TIER","GPA_HIGHSCHOOL_imp")] %>% 
  marginplot(delimiter = "_imp")
marginplot(kNN(log10(myup2020NAomit2[,c("APPLICANT_TIER","GPA_HIGHSCHOOL")])),delimiter="_imp")
imputeHS <- mice(myup2020NAomit[, c("APPLICANT_TIER","GPA_HIGHSCHOOL","ACT_PROPORTION"),m=3,seed=123])
imputeHS$imp$GPA_HIGHSCHOOL
myup2020NAomit[752:759,"ACT_PROPORTION"]
myup2020NAomitNEW <- complete(imputeHS, 4);colnames(myup2020NAomitNEW) <- c("NewTier","NewGPAHS","NewACT")
myup2020NAomitNEWOLD <- cbind(myup2020NAomit,myup2020NAomitNEW);glimpse(myup2020NAomitNEWOLD)
```


