---
title: "R Notebook_shaping"
output:
  html_document:
    df_print: paged
---


```{r}
library(data.table)
library(ISLR)
library(dplyr)
```

```{r}
data("College")
str(College)
dtcollege <- data.table(College, keep.rownames = TRUE)
mysubset <- dtcollege[40:60,];colnames(mysubset);class(mysubset)
round(mean(mysubset$Enroll), digits = 1)
```
```{r}
#extract some values in data table
x <- dtcollege[!(Outstate) %between% c(8000,14000)];x
x1 <- dtcollege[!(rn)=="Claflin College"];x1

```
```{r}
 head(dtcollege[order(-Top10perc), rn], 10) # from to10%HS by name high to low
 head(dtcollege[order(Grad.Rate
), rn],10) # for s.fRation by low to high
```

```{r}
# created a new colname
dtcollege$EARatio <-(dtcollege$Enroll/dtcollege$Accept)

#order with EARatio, Enroll and Accept as table
head(dtcollege[order(-EARatio), .(rn, EARatio, Enroll, Accept)], 10)

head(dtcollege[order(-S.F.Ratio) & Room.Board < 4000, rn], 10)
```
```{r}
subsetforunder4000 <- dtcollege[Room.Board<4000]
head(subsetforunder4000[order(S.F.Ratio), .(rn, S.F.Ratio)],10) # low to high
```
```{r}
  
# select columns
college.gradnr <- dtcollege[, c("rn","Apps","Accept","Enroll","F.Undergrad","P.Undergrad")];college.gradnr
```
```{r}
# create colname indicate highinterest for apps > 10000
dtcollege[Apps> 10000, HIGHINTEREST := 1L];tail(dtcollege,10) # := means create col and 1L means 1 or NA
# to delete the col
dtcollege[,c("HIGHINTEREST") := NULL];tail(dtcollege)
```
```{r}
head(dtcollege[ ,undergradration := F.Undergrad/P.Undergrad][order(-undergradration)], 10)
```
```{r}
#how many
dtcollege[Expend>20000, .N]
dtcollege[Books + Room.Board > 7000, .N]
dtcollege[ ,  by=Private, .N]
#how many by combination
#setDT(dtcollege)[, .N, by=.(rn,Private)]

```
```{r}
#set key words
setkey(dtcollege , F.Undergrad, P.Undergrad);head(dtcollege) #give you increasing order by F.Under

#filter
dtcollege[ F.Undergrad < 1000 & P.Undergrad < 100]
# point for 393 F and 4 P
dtcollege[ .(393,4)]
```
```{r}
# get a data.table for some colms
dtcollege[, !c("Apps","Accept","Enroll"), with=F] #false give number of cols
#keep the cols
dtcollege[ , 3:5, with=F]
```
```{r}
#how many colleges with <800 applic received have a Top10perd > 40? two conditions
dtcollege[ Apps< 800 & Top10perc > 40 , .N]

#How many with three conditions
dtcollege[Apps < 900 & Outstate<10000 & Top10perc>30, .N]

#How many three conditions and or
dtcollege[ Apps < 1000 & (Top10perc>20 | Top25perc>30), .N]
# other way
dtcollege[ Apps < 1000, .N , by = (Top10perc>20 | Top25perc>30)] # give you total apps and break part by

```
### dplyr
```{r}
library(dplyr)
EW_aveAPR_FTIC2020_lists_all <- read.csv("~/DataShaping/EW_aveAPR_FTIC2020_lists_all.csv", stringsAsFactors=TRUE)


 mydf <- data.frame(EW_aveAPR_FTIC2020_lists_all)
head(mydf)
colnames(mydf)
arrange(mydf, -BelowGPA)
slice(mydf, c(1:10))
colnames(mydf)
rename_(mydf, "Tier" = "APPLICANT_TIER") #worked
distinct_(select(mydf, ENTRY_PROGRAM))
head(mutate(mydf, BelowPr = BelowGPA*100),10)
sample_n(mydf, 10)
slice(mydf, c(1:5)) # row IDS
```

```{r}
# mutating joins column levels
testdf1 <- read.csv("~/Datashaping/CRSCSEdata2020.csv")
testdf2 <- read.csv("~/DataShaping/CSEdata2020.csv")
JI <- inner_join(testdf1,testdf2, by=c("UNIV_ROW_ID"="UNIV_ROW_ID"))
JL <- left_join(testdf1, testdf2)

#filtering joins
JA <- anti_join(testdf1,testdf2)#uniquerows for both dataset (filter on A dataset)
JS <- semi_join(testdf1, testdf2)#common row for both dataset (filter on A dataset)

#set operations

#union(testdf1,testdf2) # returens all data, but identical entries show up onl once in the resutls:complementary
#intersect(testdf1,testdf2) # returens only the identical entries and exludes everthing else; complementary
#setdiff(testdf1,testdf2)#returns the data of x, but exclude the entries which are also present in Y
```
```{r}
library(dbplyr)
#translating R to SQL
translate_sql(5) #numeric
translate_sql(5L) # L for integer in R
translate_sql(x) # object
translate_sql("x") # string
translate_sql(x-500)
translate_sql(sin(x)-500) # math function
translate_sql(x^2 < 400)
translate_sql(x ==1 && (t<100 || w>300)) # boolean
translate_sql(x %% 2 == 10) # modulo division
translate_sql(if ( x > 11) "new" else "old") # if statements
translate_sql(x < 11 & !(y>=2)) # boolean


```
```{r}
# using SQL on datasets in R
library(sqldf)
library(ggplot2)
colnames(EW_aveAPR_FTIC2020_lists_all)
sqldf("SELECT * FROM EW_aveAPR_FTIC2020_lists_all WHERE BelowGPA > 0.5")
sqldf("SELECT * FROM EW_aveAPR_FTIC2020_lists_all WHERE AboveGPA < 0.5")
sqldf("SELECT * FROM EW_aveAPR_FTIC2020_lists_all WHERE AlertR1 = 'NoEW' AND (BelowGPA > 0.5 OR aveAPR = 'Non-aveAPR')")
sqldf("SELECT * FROM EW_aveAPR_FTIC2020_lists_all WHERE EWNum >1")

```

```{r not to run}
# Get the SQLite 
#install.packages("RSQLite")
library(RSQLite)
#mydfresults <- rbind.data.frame(results12020, results22020,results32020,results42020,results52020)
#mydfresults <- EW_aveAPR_FTIC2020_lists_all
#empty database
#connection <- dbConnect(RSQLite::SQLite(), 'newmyresults1.db')
#dbWriteTable(connection, "newmyresults1", mydfresults) #write table connection new and old
#dbListTables(connection) # call newmyresults
#dbListFields(connection, "newmyresults1") # all variables from old
#dbGetQuery(connection, "SELECT * FROM newmyresults1 WHERE BelowGPA > 0.5")

```
# Set up database in r using waring data
```{r}
install.packages("sqldf")
library(sqldf)
```
# Outliers in data set
## remove / leave them / replace with mean or different values
### three sigma edit rule mean +- t*SD (ESD method)
### boxplot method Q1-c*IQD & Q3-c*IQD, IQD=Q3-Q1, c=1.5
### advanced methods for multivariate data model based method or prosimity based method
### mark ouliers with different colors or shapes - visualization- boxplot, scatterplot
## using library( mvoutlier) in R
## use a stats tests + distanc emeasure approach : ESD mehtod
```{r}
set.seed(123)
x = c( rnorm(10), 150);x
t=3
m= mean(x)
s=sd(x)
b1= m-t*s;b1
b2= m+t*s;b2
y= ifelse(x>= b1 & x<= b2, 0,1);y
plot(x, col=y+10)
```
## boxplot method
```{r}
boxplot(x)
boxplot.stats(x)
```
## package outliers
```{r}
library(outliers)
dixon.test(x)
grubbs.test(x, type=11, two.sided = T) # type=11 fpr a test for tw0 outliers on opposite tails.

```
## advaced techniques for multivariate data
```{r}
library(dplyr)
EW_all <- read.csv("~/DataShaping/EW_aveAPR_FTIC2020_lists_all.csv", stringsAsFactors=TRUE)
MAC1147EW <- EW_all %>% 
  filter(Course.Number.x == "MAC1147")
EW1147 <- MAC1147EW %>% 
  group_by(Section.Name) %>% 
  summarise(cFTIC=n());EW1147
```
```{r}
NBC_By_Tier_resultsUWF <- read.csv("~/Data_IR/NBC_By_Tier_resultsUWF.csv", stringsAsFactors=TRUE)
myoutdf <- NBC_By_Tier_resultsUWF %>% 
  filter(COHORT_YEAR< 20202021) %>% 
  select(GPA_HIGHSCHOOL,HOURS_BROUGHT_TO_UNIVERSITY,APPLICANT_TIER,CRSNAME1) %>% 
   na.omit();myoutdf

```

```{r}
library(mvoutlier)
myout <- sign1(myoutdf[, 1:3], qqcrit=0.975)#;myout #$wfinal01
myout = pcout(myoutdf[,1:3])
plot(myoutdf$GPA_HIGHSCHOOL, myoutdf$HOURS_BROUGHT_TO_UNIVERSITY, col=myout$wfinal01+2)
myoutdf$out <- myout$wfinal01
glimpse(myoutdf)
notoutliers <- myoutdf %>% 
  #filter(out == 1) %>%
  group_by(APPLICANT_TIER, out) %>% 
  summarise(meanHS=mean(GPA_HIGHSCHOOL),SDHS=sd(GPA_HIGHSCHOOL),maxHS=max(GPA_HIGHSCHOOL),minHS=min(GPA_HIGHSCHOOL),
            meanPrior =mean(HOURS_BROUGHT_TO_UNIVERSITY),sdPrior=sd(HOURS_BROUGHT_TO_UNIVERSITY), maxPrio=max(HOURS_BROUGHT_TO_UNIVERSITY),
            minPrio=min(HOURS_BROUGHT_TO_UNIVERSITY), CFTIC=n());notoutliers
```

```{r}
library(stringr)
# no numbers in string
myall <- as.data.frame(myoutdf);class(myall)
nonumner <- gsub("\\d","", myall$CRSNAME1[1:10]);head(nonumner,10)
onlynumber <- gsub("\\D","", myall$CRSNAME1[1:10]);head(onlynumber,10)
class(nonumner)
str_count(myall$CRSNAME1[1:10],"ENC1101")
all$new <- str_c(c(myall$CRSNAME1, myall$CRSNAME1), sep="");head(all$new,10) # not conc

avoidspace <- gsub("\\s","", myoutdf$CRSNAME1);head(avoidspace,10)
exchange <- gsub("[MAC]","MATH",myoutdf$CRSNAME1);head(exchange,10)
```
```{r}
glimpse(myoutdf)

```

