---
title: "Early Warning (EW) Rate and Lab Schedule - MAC1147(Lab) Fall 2020"
author: "Jay Sparks"
date: "10/29/2020"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(mosaic)   # Load additional packages here 
library(dplyr)
library(Hmisc)
library(ggplot2)
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="large")   # slightly smaller font for code
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output
#### Find the early warning rates using numbers of registerd FTIC and numbers of EW FTIC 
#### and determine the effect of the EW rates.


1. Data description
  + FTIC 2020 
  + Lists of courses of EW (check point 2)
  + Section numbers
  + Numbers of FTIC for each section
  + Numbers of EW FTIC for each section

## Top 20 EW courses for Fall 2020

```{r cars, echo = TRUE}
EWall_lists_courses <- read.csv("G:/My Drive/EW/EW_aveAPR_FTIC2020_lists_all.csv", stringsAsFactors=TRUE) %>%
   filter(AlertR1 != "NoEW" ) %>% 
  group_by(Course.Number.x  ) %>% 
  summarise(Count_EW=n(), .groups='drop') %>%
  ungroup() %>% 
  arrange(-Count_EW)
EWtop20 <- EWall_lists_courses[1:20,]

knitr::kable(EWtop20)
```

## Slide with Plot

```{r pressure}
MAC1147 <- read.csv("G:/My Drive/EW/EW_aveAPR_FTIC2020_lists_all.csv", stringsAsFactors=TRUE) %>%
   filter(Course.Number.x == "MAC1147" ) %>% 
  group_by(Course.Number.x,Section.Name) %>% 
  summarise(Count_EWFTIC=n(), .groups='drop') %>%
  ungroup() %>% 
  arrange(Section.Name)
knitr::kable(MAC1147)
```
```{r}
MAC1147$Registered_Number <- c(20, 18,21,19,21,15,21,21,21,21,22,22,18,10,23,23)
MAC1147$Lab_Day <- factor(c("FRI","FRI","FRI","FRI","FRI","FRI","FRI","FRI","FRI","FRI","FRI","THU","FRI","FRI","THU","THU"))
MAC1147$Lab_Star_Time <- factor(c("8am","10am","12:30pm","2:30pm","8am","10am","12:30pm","3pm","8am","10am","12:30pm","2:30pm","8am","2pm","11:30am","1:45pm"))
MAC1147$AMPM <- factor(c("AM","AM","PM","PM","AM","AM","PM","PM","AM","AM","PM","PM","AM","PM","AM","PM"))
MAC1147$Rate_EW <- round((MAC1147$Count_EWFTIC/MAC1147$Registered_Number), digits = 4)
head(MAC1147[,2:8])
```
```{r}
summary(MAC1147)
Total_Registed <- sum(MAC1147$Registered_Number);Total_Registed
```
```{r}

ggplot(data=MAC1147, aes(x=Registered_Number, y= Rate_EW, color= Lab_Day)) + geom_point(size=3)+
  labs(title= "Early Warning Rate and FTIC 2020",
       subtitle = "Source IR/EW 2020")
```
```{r}
MAC1147rate <- MAC1147 %>% 
  group_by(Lab_Day) %>% summarise(meanEW_Rate = mean(Rate_EW), .groups="drop");MAC1147rate
```

### Find a model to predict the EW rate using size of class, lab times, and days
###The results suggest that the Thursday lab has a significant effect on the EW rate.
###The p-value is 0.1898658 and more details are below.
```{r}
library(broom)
model1 <- lm(Rate_EW ~ Registered_Number + Lab_Star_Time + Lab_Day , data= MAC1147);mtab <- tidy(model1)
knitr::kable(mtab)
```
