---
title: "R Notebook_Advanced  programming"
output:
  html_document:
    df_print: paged
---
### import data
[The data is in this website] (https://www.superdatascience.com/pages/rcourse-advanced)
```{r}
getwd()
setwd("G:/My Drive/R_learning/Repeated_F")
P3.Future.500.The.Dataset <- read.csv("G:/My Drive/R_learning/Repeated_F/P3-Future-500-The-Dataset.csv",
                                      stringsAsFactors = T, na.strings = c(""))
fin <- P3.Future.500.The.Dataset
fin$ID <- factor(fin$ID)
str(fin)
summary(fin)
```
### convert variables factor to characters and numberic
```{r}
typeof(fin$ID)
b <- as.character(fin$ID)
typeof(b)
c <- as.numeric(b)
typeof(c)

fin$Profit <- as.character(fin$Profit)
fin$Profit <- as.numeric(fin$Profit)
typeof(fin$Profit)
str(fin)
summary(fin)
```
###  gsub
```{r}
fin$Expenses <- gsub("Dollars", "", fin$Expenses );head(fin,10)
fin$Growth <- gsub("%", "", fin$Growth);head(fin, 10)
fin$Revenue <- gsub("\\$","", fin$Revenue);head(fin,10)
fin$Revenue <- gsub(",", "", fin$Revenue)
fin$Expenses <- gsub(",", "", fin$Expenses)
fin$Revenue <- as.numeric(fin$Revenue)
fin$Growth <- as.numeric(fin$Growth)
fin$Expenses <- as.numeric(fin$Expenses);str(fin)
```
### look up excel sheet how we can replace missing data 
### some variables can be proxy not all such as year or unique ID name
### TRUE 1 and FLASE 0 and NA
```{r}
head(fin,24)
get.all.with.NA <- fin[!complete.cases(fin),];get.all.with.NA
```
### using which() for filtering
```{r}
fin[fin$Revenue==9254614,]
filter.with.na <- fin[fin$Revenue==9254614,];filter.with.na
filter.row.numer <- fin[which(fin$Revenue==9254614),];filter.row.numer
filter.row.numer1 <- fin[which(fin$Industry=="Health"),];filter.row.numer1
```
###filtering using is.na()
```{r}
med_exp_constr <- median(fin[fin$Industry== "Construction","Expenses"], na.rm = T);med_exp_constr
fin[is.na(fin$Expenses) & fin$Industry == "Construction" & is.na(fin$Profit),]
fin[is.na(fin$Expenses) & fin$Industry == "Construction" & is.na(fin$Profit), "Expenses"] <- med_exp_constr
fin[is.na(fin$Profit),"Profit"] <- fin[is.na(fin$Profit),"Revenue"] - fin[is.na(fin$Profit),"Expenses"]
fin[c(8,42),]
```
##ggplot
```{r}
library(ggplot2)
p <- ggplot(data=fin, aes(x=Revenue, y= Expenses, colour=Industry))+geom_point()+geom_smooth(fill=NA, size=1.2)
p
```

```{r}
p1 <- ggplot(data=fin, aes(x=Industry, y= Growth, colour=Industry))+geom_jitter()+geom_boxplot(size=1, alpha=0.5, outlier.color = NA) 
p1

```
##import data name machine utilization
```{r}
util <-read.csv("G:/My Drive/R_learning/Repeated_F/P3-Machine-Utilization.csv")
head(util,20)
```
```{r}
util$Utilization=1-util$Percent.Idle
head(util,20)
```
## handing data to time
```{r}
tail(util,10)
```
## POSIXlt POSIXct (portable  operating system inerface )
```{r}
util$Posixtime <- as.POSIXct(util$Timestamp, format="%d/%m/%Y %H:%M");head(util,10)
```
## how to rearrange column in a df:
```{r}
util$Timestamp <- NULL
head(util,10)
util <- util[ , c(4,1,2,3)];head(util,10)
```
## What is a list?
```{r}
util$Machine <- factor(util$Machine)
summary(util)
RL1 <- util[util$Machine=="RL1",]
summary(RL1)
RL1$Machine <- factor(RL1$Machine)
summary(RL1)
```
## Get stat with one machine with below 90% utilization as list
```{r}
util_stats_RL1 <- c(min(RL1$Utilization, na.rm=T),mean(RL1$Utilization, na.rm=T) ,max(RL1$Utilization, na.rm=T));util_stats_RL1
util.under90.flag.logic <-  length(which(RL1$Utilization < 0.9,)) > 0 ; util.under90.flag.logic
list.rl1 <- list("RL1", util_stats_RL1,util.under90.flag.logic);list.rl1
```
## components of a list

```{r}

list.rl1
names(list.rl1)
names(list.rl1) <- c("Machine","stats","LowThreshold")
list.rl1

list.rl1.names <- list(Machine="RL1", Stats=util_stats_RL1,LowThreshold=util.under90.flag.logic);list.rl1.names


```
## extracting components of a list
### three ways: [] return a list, [[]] return the actural object, and $ same as [[]]
```{r}
list.rl1.names[1]
typeof(list.rl1.names$Machine)
list.rl1.names[[2]][3]
list.rl1.names$Stats[3]
```
## adding and deleting list components

```{r}

list.rl1.names[7] <- "New.Information"
list.rl1.names
list.rl1.names$unknownTime <- RL1[is.na(RL1$Utilization),"Posixtime"]
list.rl1.names
list.rl1.names$Data <- RL1$Posixtime
list.rl1.names[[8]] <- NULL
summary(list.rl1.names)

list.rl1.names[[5]][1]
list.rl1.names[1:4]
sub.list <- list.rl1.names[c(1,4,7)]
summary(sub.list)
```

## build time series plot and create list with plot to organize your work and space
```{r}
library(ggplot2)
p <- ggplot(util)
myutiltimeplot <- p+ geom_line(aes(x=Posixtime, y=Utilization, colour=Machine), size=1.1) +
  facet_grid(Machine~.) +
  geom_hline(yintercept = 0.90, colour="gray", size=1.1, linetype=2)
list.rl1.names$Plot <- myutiltimeplot
list.rl1.names
str(list.rl1.names)
```
## "Apply" family of functions: shortcut for loop
### import data and convert first row as names
### convert data.frame as matrix
```{r  knitr root.dir }
getwd()
setwd("./Weather Data")
getwd()
Chicago <- read.csv("Chicago-F.csv", row.names = 1)
NewYork <- read.csv("NewYork-F.csv", row.names = 1)
Houston <- read.csv("Houston-F.csv", row.names = 1)
Sanfrancisco <- read.csv("Sanfrancisco-F.csv", row.names = 1)
is.data.frame(Chicago)


Chicago <- as.matrix(Chicago)
NewYork <- as.matrix(NewYork)
Houston <- as.matrix(Houston)
Sanfrancisco <- as.matrix(Sanfrancisco)

is.matrix(Chicago)
```
## put the matrix as list
```{r}
Weather <- list(Chicago=Chicago, NewYork=NewYork, Houston=Houston, Sanfrancisco=Sanfrancisco)
Weather[[3]]
Weather[[3]][3]


```
## What is apply function in r
### apply(matrix, 1, mean) row loop, apply(M,2,mean) column loop
### tapply use on a vector to extreact sugroyps
### by use data frame in group by in SQL
### eapply use on an environment(E)
### lapppy apply to lelement sof list
### sapply a version of lapply can simplyfiy not as list
### vapply returne values
### replicate run a function several time random varibles
### mapply mltivarite version of sapply and arguments can be recycled
### rapply recursive vversion of lapply
```{r}
apply(Chicago, 1, mean)
mean(Chicago["HoursOfSunshine",])
apply(Chicago, 1, min)
apply(Chicago, 1, max)
apply(Chicago, 1, median)

```
## recreatubg a apply functin with loops
```{r}

output <- NULL
for (i in 1:5) {
 output[i]  <- mean(Chicago[i,])
  
}
names(output) <- rownames(Chicago)
output

apply(Chicago, 1, mean)

```
### transfose data using t function and learn rowMeans, colMeans, rowSums,colSums

```{r}
t(Chicago)

newlistweaterh <- lapply(Weather, t)
newlistweaterh

newlistweahter2 <- lapply(Weather, rbind, NewRow=1:12)
newlistweahter2

newlistweaater3 <- rowMeans(Chicago)
newlistweaater3
newlistweather4 <- lapply(Weather, rowMeans)
newlistweather4

lapply(Weather, rowSums)
Weather$Chicago
```
## combining lapply "[" means []

```{r}
Weather[[1]][1,1]
lapply(Weather, "[", 1,)
lapply(Weather, "[", 1,2)
lapply(Weather, "[", )



```

## Adding own function
```{r}
Weather
Weather$maxmindiff <- lapply(Weather, function(x) round((x[1,]-x[2,])/x[2,],2))
Weather$maxmindiff <- NULL
```

```{r}
str(Weather)
lapply(Weather, "[", 2, 7:12) #chicago July
round(sapply(Weather, rowMeans),2)
sapply(Weather, rowMeans, simplify = F)

```

## Nesting Apply function
```{r}
lapply(Weather, apply, 1, max)
lapply(Weather, function(x) apply(x, 1, max))

sapply(Weather, function(x) apply(x, 1, max))
sapply(Weather, function(x) apply(x, 1, min))
sapply(Weather, function(x) apply(x, 2, mean))


Weather
```
## Which.max/ which.min/ by components by whole lists
```{r}
which.max(Chicago[1,])
names(which.max(Chicago[1,]))

apply(Chicago, 1, function(x) names(which.max(x)))

lapply(Weather, function(y) apply(y, 1, function(x) names(which.max(x))))
sapply(Weather, function(y) apply(y, 1, function(x) names(which.max(x))))

```




