---
title: "Who is most likely at risk"
author: "JS"
date: "2/10/2021"
output:
  powerpoint_presentation: default
  beamer_presentation: default
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

```{r include=FALSE}
#################################################
#### Association Rule for prior hours == 0 ######
#################################################


cleanFTIC_all_falls_BdegDF <- read.csv("~/lowerGPA_all_falls/cleanFTIC_all_falls_BdegDF.csv", stringsAsFactors=TRUE)

xtabs(~cleanFTIC_all_falls_BdegDF$Cohort+cleanFTIC_all_falls_BdegDF$Stu_FirstGenInd)
xtabs(~cleanFTIC_all_falls_BdegDF$Cohort+cleanFTIC_all_falls_BdegDF$Stu_Ethnicity)
xtabs(~  Cohort+cleanFTIC_all_falls_BdegDF$Deg_College+cleanFTIC_all_falls_BdegDF$Stu_Ethnicity,data=cleanFTIC_all_falls_BdegDF)



FTIC_GPA_com <- cleanFTIC_all_falls_BdegDF %>% 
  group_by(Termindex,PriorUniAPR,APRGPA1stFall) %>% 
  summarise(meanGPA = mean(GPAEachFalls),meanTermHrs=mean(Stu_GPATermHours),meanPriorHrs=mean(Stu_TotalUniversityHours),CountFTIC=n())  #(meanGPA = mean(GPA1stFall),
FTIC_GPA_com


summary(cleanFTIC_all_falls_BdegDF)


############################################
#### select variables  #####################
###########################################

glimpse(cleanFTIC_all_falls_BdegDF)

outDEGGPAFTIC_1st_term_association <- cleanFTIC_all_falls_BdegDF %>% 
  filter( PriorUniAPR =="NoPriorHrs" & Termindex== 1) %>% # for the 1st fall
  filter(GPAEachFalls >  0.28626667 , na.rm=TRUE) %>%  # outliers <=  0.2866667
  mutate(codeAGE=ifelse(Stu_Age<=18,"Age<=18","Age>18")) %>% 
  mutate(TargetedDegreeArea=ifelse(Stu_StrategicEmphasis=="STEM","STEM",ifelse(Stu_StrategicEmphasis=="Health","Health","nonSTEM/Health"))) %>% 
  select(Gender="Stu_Gender",County="Stu_County",Ethnicity="Stu_Ethnicity",Enter_college="Stu_College",
         FirstGeneration="Stu_FirstGenInd",codeAGE,TargetedDegreeArea,
         ResidenceHall="Stu_ResidenceHall",
         #Stu_GPATermHours,
        #rankGPAEachFall,
        FirstFallGPA="APRGPA1stFall",
        #Deg_TermGranted
        #,DegCollege="Deg_College"
        ) #1490/1617
FOR_belowFTIC <- outDEGGPAFTIC_1st_term_association %>% filter(FirstFallGPA=="Below2.77") %>% select(-9)

summary(outDEGGPAFTIC_1st_term_association)

#check NA
p <- function(x){sum(is.na(x))/length(x)*100}
apply(outDEGGPAFTIC_1st_term_association, 2, p)
table(outDEGGPAFTIC_1st_term_association$County)
```

```{r}


# need libraries
library(arules)
library(arulesViz)


#summary
DEGGPAFTIC_1st_term_associationDF <- as(outDEGGPAFTIC_1st_term_association,"transactions") #transaction are analysed to identify rules of assocation
class(DEGGPAFTIC_1st_term_associationDF)

FOR_belowFTIC_transactions <- as(FOR_belowFTIC, "transactions")
#inspect(head(newGPAassociationDF)) # too long
fig_1 <- arules::itemFrequencyPlot(DEGGPAFTIC_1st_term_associationDF, supp=0.15,  
                          col = brewer.pal(8, 'Pastel2'), 
                          main = 'More Than 15% Levels Frequency Plot - Inexperienced FTIC', 
                          type = "relative", 
                          ylab = "Factor Frequency(Relative)")
```

```{r}
arules::itemFrequencyPlot(DEGGPAFTIC_1st_term_associationDF, topN=20,  
                          col = brewer.pal(8, 'Pastel2'), 
                          main = 'Top 20 Frequency Plot - Inexperienced FTIC', 
                          type = "absolute", 
                          ylab = "Factor Frequency (absolute)")

```

```{r}
FOR_belowFTIC_transactions <- as(FOR_belowFTIC, "transactions")
#inspect(head(newGPAassociationDF)) # too long
arules::itemFrequencyPlot(FOR_belowFTIC_transactions, supp=0.01,  
                          col = brewer.pal(8, 'Pastel2'), 
                          main = 'More Than 10% Levels Frequency Plot - Underperformed & Inexperienced FTIC', 
                          type = "relative", 
                          ylab = "Factor Frequency(Relative)")
```



```{r}
# results for above2.7
rules_GPAabove <- apriori(DEGGPAFTIC_1st_term_associationDF,parameter = list(minlen=1, maxlen=6, conf=.8, support=0.01), 
                          # to see whether there is sufficient evidence to suggest
                    appearance = list(rhs=c("FirstFallGPA=Above2.77"), default = "lhs"))

sum(is.redundant(rules_GPAabove)) #

above2.77 <- rules_GPAabove[!is.redundant(rules_GPAabove)]
resultsabove <- inspectDT(head(sort(above2.77, by="lift"),30)) 
# three metrics (most commonly)
#support:%of transactions that contain all of the items in an itemsets, the higher values the more frequently the itemset occurs
# rules with high support are preferred since more likely to be applicable to a large number of future transactions
#confidence:probability that lhs of the rule also contain rhs
#lift: strengh of association
#coverage: lhs support

resultsabove
outputabove2.77 <- cbind(label=labels(above2.77), quality(above2.77))
table_above <- DT::datatable(outputabove2.77)
write.table(outputabove2.77,"Results_RiskFTIC_above.csv",sep=",")


#save as html
htmlwidgets::saveWidget(resultsabove, "arulesabove.html", selfcontained = FALSE)
browseURL("arules.html")
```


```{r}
### Below2.77
rules_GPAbelow <- apriori(DEGGPAFTIC_1st_term_associationDF,parameter = list(minlen=1, maxlen=6, conf=.6, support=0.01), # to see whether there is sufficient evidence to suggest
                     appearance = list(rhs=c("FirstFallGPA=Below2.77"), default = "lhs"))
sum(is.redundant(rules_GPAbelow)) #
below2.77 <- rules_GPAbelow[!is.redundant(rules_GPAbelow)]
resultsbelow <- inspectDT(head(sort(below2.77, by="lift"),30)) # three metrics (most commonly)
#support:%of transactions that contain all of the items in an itemsets, the higher values the more frequently the itemset occurs
# rules with high support are preferred since more likely to be applicable to a large number of future transactions
#confidence:probability that lhs of the rule also contain rhs
#lift: strengh of association
#coverage: lhs support
resultsbelow
outputbelow2.77 <- cbind(label=labels(below2.77), quality(below2.77))
table_below <- DT::datatable(outputbelow2.77)
write.table(outputbelow2.77,"Results_RiskFTIC.csv",sep=",")
```


```{r}
### risk FTIC
FOR_belowFTIC_a <- outDEGGPAFTIC_1st_term_association %>% filter(FirstFallGPA=="Below2.77") 
FOR_belowFTIC_a_transactions <- as(FOR_belowFTIC_a, "transactions")
# to see whether there is sufficient evidence to suggest
rules_GPAbelow_risk <- apriori(FOR_belowFTIC_a_transactions,parameter = list(minlen=1, maxlen=6, conf=.9, support=0.5), 
                                               appearance = list(rhs=c("FirstFallGPA=Below2.77"), default = "lhs"))
sum(is.redundant(rules_GPAbelow_risk)) #

riskbelow2.77 <- rules_GPAbelow_risk[!is.redundant(rules_GPAbelow_risk)]
resultsbelowRisk <- inspectDT(head(sort(riskbelow2.77, by="lift"),30)) # three metrics (most commonly)
#support:%of transactions that contain all of the items in an itemsets, the higher values the more frequently the itemset occurs
# rules with high support are preferred since more likely to be applicable to a large number of future transactions
#confidence:probability that lhs of the rule also contain rhs
#lift: strengh of association
#coverage: lhs support
resultsbelowRisk
outputbelow2.77risk <- cbind(label=labels(riskbelow2.77), quality(riskbelow2.77))
table_below_risk<- DT::datatable(outputbelow2.77risk)
write.table(outputbelow2.77risk,"Results_BelowRiskFTIC.csv",sep=",")



```


```{r}
###plots
p_below <- plot(below2.77, method="graph", engine = "interactive",shading = "lift") #grouped



## find reduncdant
sum(is.redundant(rules_02))
inspect(rules_01[is.redundant(rules_02)])
newM <- DATAFRAME(rules_02)
newM[newM$RHS == "{averageGPA1stFall=Below2.7}" & grepl(x=newM$LHS, pattern = "\\{Non-Florida\\}|\\{White}"),]#??

rerules_02 <- rules_02[!is.redundant(rules_02)]

```


```{r}
##plots
require(arulesViz)
plot(x=below2.77,
     measure = c("confidence","lift"), shading = "support")

plot(x=below2.77,
     method = "scatterplot", main="Scatter plot for above GPA Rules")

plot(x=below2.77,
     method = "two-key plot") # higher order give higher confidence but support

plot(x=below2.77,
     method = "matrix")

plot(x=below2.77,
     method = "matrix3D")

plot(x=below2.77,
     method = "graph",
     engine = "htmlwidget")


```


```{R}
#crossTable
ct <- crossTable(DEGGPAFTIC_1st_term_associationDF,  sort=TRUE)
ct
sp <- crossTable(DEGGPAFTIC_1st_term_associationDF,measure="support", sort=TRUE)
sp
lift <- crossTable(DEGGPAFTIC_1st_term_associationDF, measure="lift", sort=TRUE)
lift[1:5,1:5]
chi2 <- crossTable(DEGGPAFTIC_1st_term_associationDF, measure="chiSquared", sort=TRUE)
round(chi2[1:5,1:5], digits = 5)


#################################################################################
#### Association Rule for Inexperienced FTIC using CSEdata ( for APR GPA)  ######
#################################################################################




library(readr)
CSEdata <- read_csv("~/Data_IR/CSEdata.csv", 
                    col_types = cols(UNIV_ROW_ID = col_character()))
View(CSEdata)
glimpse(CSEdata)
apply(CSEdata, 2, p)

outliersAPR <- CSEdata$GPA_ENTERING_SECOND_FALL
outL <- boxplot.stats(CSEdata$GPA_ENTERING_SECOND_FALL)$out 
outL# 0.990000 2.525735 3.106896 3.555719 4.000000
min(outL, na.rm=TRUE) #0
max(outL, na.rm = TRUE) #0.98
glimpse(CSEdata)
### Filter not outliers

CSEdataV1 <- CSEdata %>% 
  filter(GPA_ENTERING_SECOND_FALL >= 0.99, na.rm=TRUE )  %>% # remove outliers
  mutate(AverageGPAindex = ifelse(GPA_ENTERING_SECOND_FALL< 3.106896, "Below3.10","Above3.10")) %>% 
  mutate(ExpFTIC=ifelse(HOURS_BROUGHT_TO_UNIVERSITY<= 0,"InexpFTIC","ExpFTIC"))
apply(CSEdataV1, 2, p)   # no NA for GPA

FTIC_GPADEG_com <- CSEdataV1 %>% 
  group_by(COHORT_YEAR,ExpFTIC,AverageGPAindex) %>% 
  summarise(meanPriorHrs=mean(HOURS_BROUGHT_TO_UNIVERSITY),CountFTIC=n())  #(meanGPA = mean(GPA1stFall),
FTIC_GPADEG_com



#################################
######## select variables #######
#################################

CSEdataV2 <-  CSEdataV1 %>% 
  #filter(ExpFTIC=="InexpFTIC") %>% 
  select(ENTRY_COLLEGE,
         ENTRY_DEPARTMENT,
         GENDER,
         ETHNICITY,
         COUNTY,
         ENTRY_PROGRAM,
         #COUNTY,
         FIRST_GENERATION_STUDENT,
         #ATHLETE,
         APPLICANT_TIER,
         HIGH_SCHOOL_NAME,
         GPA_HIGHSCHOOL,
         HOURS_BROUGHT_TO_UNIVERSITY,
         AP_CREDITS,
         ACT_PROPORTION,
         SAT_PROPORTION,
         #FIRST_FALL_PELL,FIRST_FALL_BRIGHT_FUTURES,FIRST_FALL_ANY_LOANS,
         #FIRST_FALL_NEED_BASED_LOANS,
         AverageGPAindex)

summary(CSEdataV2)
# need libraries
library(arules)
library(arulesViz)


#summary
CSEdataV3 <- as(CSEdataV2,"transactions") #transaction are analysed to identify rules of assocation
class(CSEdataV3)

summary(CSEdataV3)
#inspect(head(newGPAassociationDF)) # too long
itemFrequencyPlot(CSEdataV3, topN=10)
itemFrequencyPlot(CSEdataV3, support=0.05)

# results for above3.10
rules_above <- apriori(CSEdataV3,parameter = list(minlen=1, maxlen=6, conf=.5, support=0.1), # to see whether there is sufficient evidence to suggest
                       appearance = list(rhs=c("AverageGPAindex=Above3.10"), default = "lhs"))
results_above <- inspect(head(sort(rules_above, by="lift"),10))
results_above


# results for below3.10
rules_below <- apriori(CSEdataV3,parameter = list(minlen=1, maxlen=6, conf=.5, support=0.2), # to see whether there is sufficient evidence to suggest
                       appearance = list(rhs=c("AverageGPAindex=Below3.10"), default = "lhs"))
results_below <- inspect(head(sort(rules_below, by="lift"),10))
results_below


#Plots
library("RColorBrewer")
plot(below2.77, control = list(brewer.pal(10,"Spectral")), main="Rules for Below average")# showing high lift(support/confidence) for rules

require(arulesViz)
plot(x=below2.77,
     measure = c("confidence","lift"), shading = "support")


```

